### 2018-04-29 2_wmm.tex
cleanup the wmm section a bit.

The relation~$\xrightarrow{r}\,:~\mathbb{E}~\times~\mathbb{E}$ is a binary function over events (set of pairs of events). There are two kinds of relations between events: \textit{basic relations} %retrieved from the 
that capture semantics of the program, and \textit{derived relations} that are defined from the basic relations and events in the weak memory model specification. Constraints over relations that are specified by weak memory models are defined as requirements of acyclicity, irreflexivity or emptiness of specific relations~\cite{alglave2016syntax}.

@@

    \item the \textit{from-read} relation $\fr \subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read to writes preceding the write event from which the read reads the value: \\
    $r \relation{fr} w = (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$.

    \item the \textit{data dependency} relation \texttt{dp}, which is a subset of \po-relation that always has a read at its source (it connects the read to the write which it depends on).
    %TODO: say that po-relation is *static*
    %$\mathtt{com} = (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$

    \item the \textit{external} (and \textit{internal}) \textit{read-from} relations that restrict the \rf-relation to the different (respectively, same) processes.

    \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable:
    $m_1 \relation{po-loc} m_2 = (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$.

    \item the semantics of fences and barriers specific for different architectures may be defined as derived relations.

@@

The work~\cite[Chapter~2]{alglave2010shared} provides definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}.
Thereafter, weak memory models make asserts over these properties, thus restricting the set of allowed behaviours of the system.

% TODO: some derived relations
%The \textit{from-read order} \fr $\subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read to possible writes preceding the current write event (this relation is the inversion of the \rf-relation: $\mathtt{rf} = \mathtt{fr}^{-1}$).
%TODO: perephrase last sentence! alglave thesis, p. 36

@@

The derived relations can be defined via the keyword \texttt{let} and the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events

>>>>

The relation~%\xrightarrow{r}
$r\,\subseteq~\mathbb{E}~\times~\mathbb{E}$ is a set of pairs of events (a subset of Cartesian product of two sets of events). There are two kinds of relations between events: \textit{basic relations} %retrieved from the 
that capture semantics of the program, and \textit{derived relations} that are defined from the basic relations and events in the weak memory model specification. Constraints over relations that are specified by weak memory models are defined as requirements of acyclicity, irreflexivity or emptiness of specific relations~\cite{alglave2016syntax}.

@@

    \item the \textit{from-read} relation $\fr \subset~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to write events preceding the write event from which the read event gets the value: \\
    $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$

    %\item the \textit{data dependency} relation \texttt{dp}, which is a subset of \po-relation that always has a read at its source (it connects the read to the write which it depends on)
    %TO-DO: formula for dp or exclude

    \item the \textit{communication} relation \com over memory events, that fully describes the data-flow of a program: \\
    %$\mathtt{com} \triangleq (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
    $m_1 \relation{com} m_2~\triangleq~((m_1\relation{rf}m_2) \lor (m_1\relation{co}m_2) \lor (m_1\relation{fr}m_2))$

    \item the \textit{external} (and \textit{internal}) \textit{read-from} relations that restrict the \rf-relation to the different (respectively, same) processes: \\
    $w \relation{fre} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) = \mathtt{proc}(w))$ \\
    $w \relation{fri} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) \neq \mathtt{proc}(w))$

    %TODO: say that po-relation is *static*
    \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable: \\
    $m_1 \relation{po-loc} m_2 ~\triangleq~ (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$

    \item the semantics of \textit{fences} (memory barriers) specific for different architectures may be defined as derived relations.

@@

The work~\cite[Chapter~2]{alglave2010shared} provides formal definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}, that can be asserted by a weak memory model in order to restrict the set of allowed behaviours of the system.

@@

New relations can be defined via the keyword \texttt{let} and the following operations over relations~\cite{alglave2016syntax}: % 3.1.6 Relations between events
